---
title: "HW2_Jingjian Wang"
author: "Jingjian Wang"
date: '2022-08-31'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# for reading csv file
library(readr) 

# for data management
library(dplyr) 

# for coding
library(tidyverse) 


```

1. Import class one survey data from our Github site (https://github.com/kijohnson/ADA) (The dataset is called ‘Class 1 Survey Fall 2022.csv’), calling the R dataframe that you create C1survey.
```{r cars}
# import dataset
C1survey<-read_csv("https://raw.githubusercontent.com/kijohnson/ADA/master/Class%201%20Survey%20Fall%202022.csv")
```
2. Determine: a. the number of observations (i.e. the number of people who filled out the survey) and b. the number of variables in the dataframe.

```{r}
# the number of observations in the dataframe
dim(C1survey)[1]
# the number of variables in the dataframe
dim(C1survey)[2]
```
3. a. Rename the column variables to something shorter and that is descriptive of what the variable is about (for example like_dogs for the ‘Do you like dogs?’ question variable) and b. Display that they are renamed.

```{r}
#renames variables in a shorter form
names(C1survey)[2:27]<-c("like_cat", "like_dog", "desert_item", "life_slogan",  "favorite_weekday", "circadian_rhythm",  "favorite_food", "favorite_drink", "favorite_season", "favorite_month", "hobby", "program", "specialization", "pri_sta_software", "level_r_exp", "level_comfort_coding", "how_long_coding", "three_things_to_learn", "interest_area", "favorite_number", "birth_day", "birth_month", "home_country", "home_state", "home_city", "level_education") 

colnames(C1survey) 
```

4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the C1survey dataset.
```{r}
# number of factor
sum(sapply(C1survey,is.factor))
# number of integer
sum(sapply(C1survey,is.integer))
# number of numeric
sum(sapply(C1survey,is.numeric))
# number of character
sum(sapply(C1survey,is.character))
```

5. a. Using code, check the bday and bmonth variables for any unusual or missing values. If you find any, b. describe what you will do with the unusual values in a comment before or immediately following the code that does it, and c. after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r}
# check for missing values and unusual values - check character values by human
C1survey <- C1survey %>%
  mutate(birth_day = recode(birth_day, "June 8th" = "8",
                                      "13-Aug-97" = "13")) %>%
  mutate(birth_month = recode(birth_month, "June 8th, 1998" = "6",
                                      "August" = "8",
                                      "October" = "10",
                                      "March" = "3")) %>%
  mutate(bday = as.numeric(birth_day)) %>%
  mutate(bmonth = as.numeric(birth_month)) %>%
# check unusual month
  mutate(month_err=if_else(1<=bmonth & bmonth<=12, 0, 1)) %>%
# check unusual day
  mutate(day_err=if_else((bmonth == 2 & bday<=29) |
                         ((bmonth == 1 | 3 | 5 | 7 | 8 | 10 | 12 ) & bday<=31) |
                         ((bmonth == 4 | 6 | 9 | 11 ) & bday<=30) , 0, if_else(bday>0 & bday<32, 0, 1)))
summary(C1survey$month_err)
summary(C1survey$day_err)

colSums(is.na(C1survey))

# median
median(C1survey$bday, na.rm = TRUE)
median(C1survey$bmonth, na.rm = TRUE)
```

6. a. Create a new variable called bseason that gives the season according to Northern Meteorological season in which respondents were born (winter=12/1 to 2/29, spring=3/1 to 5/31, summer=6/1 to 8/31, fall=9/1 to 11/30). b. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct c. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code.

```{r}
C1survey <- C1survey %>%
  mutate(bseason = recode_factor(bmonth, "3" = "Spring",
                                         "4" = "Spring",
                                         "5" = "Spring",
                                         "6" = "Summer",
                                         "7" = "Summer",
                                         "8" = "Summer",
                                         "9" = "Fall",
                                         "10" = "Fall",
                                         "11" = "Fall",
                                         "12" = "Winter",
                                         "1" = "Winter",
                                         "2" = "Winter"))
# use table to check coding results
table_season <- table(C1survey$bmonth,C1survey$bseason)
# add margins
addmargins(table_season, 1, FUN = sum)
# 7 in Spring, 11 in Summer, 10 in Fall and 9 in Winter
```

7. Pick your favorite variable to analyze, come up with a question you want to answer with that variable, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #’s).

```{r}
# put the number of observations in nobs
nobs<-as.numeric(C1survey %>% summarise(nobs=n()))
# variable circadian_rhythm
# which kind has the most - owl, lark, or hummingbird?
# generate a table
table(C1survey$circadian_rhythm)
# draw a plot to visualize
C1survey %>%
  ggplot(aes(x = circadian_rhythm)) +
  geom_bar() +
  scale_y_continuous(breaks = seq(0, 40, 2), sec.axis = sec_axis(~. / nobs * 100, name = "Percentage", breaks = seq(0, 100, 5))) +
  coord_flip() +
  theme_bw() +
  theme(plot.title.position = 'plot', plot.title = element_text(hjust = 0.5)) +
  labs(x = "Kind of bird", y = "Count", title = "Circadian Rhythm")
# most people are hummingbirds, followed by owls, only 9 people are larks
```

